% !TeX root = ../arara-manual.tex
\chapter*{Foreword}
\label{chap:foreword}

\epigraph{That deserves no less than a ``Holy guacamole!''.}{\textsc{Gonzalo Medina}}

{\setlength{\parskip}{1em}
Creating a PDF from \LaTeX\ code can be quite tiresome. Suppose I am using \TeX works and I have a document that has a bibliography, glossary and index, then I need to select the \rbox{pdflatex} tool and click on the typeset button, then select the \rbox{bibtex} tool and click on the typeset button, then select the \rbox{makeindex} tool and click on the typeset button, then select the \rbox{makeglossaries} tool (which I may need to add first) and click on the typeset button, then select the \rbox{pdflatex} tool and click on the typeset button, and once more to ensure all the cross-references are up to date. Then I edit the document and have to go through that whole process all over again!

Automation makes life so much simpler. Instead of all those tools that I need to keep selecting, I just need one tool, in this case \arara, which will do all the necessary work for me behind the scenes.

Some automation tools try to be clever, but there are invariably exceptions that trip them up. \arara\ does not try to be clever; it just does what it is told to do. The instructions are provided as special comments in the source code that \TeX\ ignores, but they are human-readable and can also provide a hint to non-\arara\ co-authors as to what tools are required in order to complete the document build.

The new improved \arara\ version 4.0 now comes with some exciting features, such as the ability to use conditionals, and it definitely ranks as my favourite automation tool for document creation. Paulo has done a great job, and I would like to take this opportunity to thank
him for his patience in dealing with my many feature requests!}

\vfill

\begin{flushright}
Nicola Louise Cecilia Talbot\\
\emph{on behalf of the \arara\ team}
\end{flushright}
